### 2011-12-27 joint action and knowing others minds/interacting mindreaders.2.tex
second section drafted, also some work on conclusion (incomplete)

Because goal ascription has received little attention,
we shall briefly review some potential benefits of being able to identify goals.
It is a familiar idea that goal ascription enables one to learn from others' successes.
For example,
if you know or can guess that another agent's actions are directed to opening a nut,
you may then be in a position to infer that the unfamiliar pattern of actions she is performing constitute a means to open nuts.
%\citep{Horner:2005pj}
In a different case, knowing the goal of another agent's actions may enable you to discover a new use for a tool.
A slightly less familiar idea is that goal ascription enables one to learn from others' failures as well as their successes.
For example, suppose that while you are searching for some peanuts 
another agent attempts but fails to reach for a closed container.
In some circumstances,
if you know that the goal of the agent's action was to obtain the peanuts
then you now have evidence as to where they might be.\footnote{
\citet{hare_chimpanzees_2004} exploit this principle in testing chimpanzees' abilities to ascribe goals.
}
This is one illustration of how goal ascription could in principle enable us to learn from others' failures 
(\citealp{Want:2001hp} offer another).

@@

Now that we have reviewed some of the benefits goal ascription can bring,
what evidence grounds ascriptions of goals to actions?
Consider the claim that 
knowledgeably identifying the goals to which actions are directed
depends on 
knowledge of the agents' intentions
which in turn depends 
on knowledge of their beliefs, desires and other mental states.
If this claim were true,
there would be no such thing as a theory of the evidential basis of goal ascription,
 only a theory of the evidential basis of mindreading more generally.

The falsity of this claim seems to be assumed in developmental research on goal ascription.%
\footnote{ 
Compare \citet{Gergely:1995sq}
and 
\citet{Woodward:1998dm}
among many others.
}
We know of no compelling reason for taking this claim to be true.
After all,
it is consistent with rejecting the claim to recognize that
 ascriptions of goals to actions
 constrain, and are constrained by,
 ascriptions of intentions and other mental states.
The existence of such constraints does not show that one could 
not know something about the goals of actions 
while not knowing anything about the agents' mental states.
Further,
there do seem to be situations where knowledge of agents' goals does not require any knowledge of their mental states.
For instance,
suppose that two people are sitting opposite each other at a low table
 which is 
sparsely populated with objects.
The objects are all out in the open; manifestly, both can clearly see them.
If one person reaches to grasp one of these objects (the duck, say), 
must the other ascribe beliefs or other mental states in order to knowledgeably identify the goal of her action as that of grasping the duck?
On the face of it, she need not.  
Even if she had no ability to ascribe mental states, it seems she might nevertheless be in a position to acquire knowledge of the goal of the other's action.

While not decisive,
these considerations are perhaps sufficient to motivate 
exploring what evidence might ground goal ascription.
According to 
what Csibra and Gergely call `the principle of rational action':
%
\begin{quote}
`an action can be explained by a goal state if, and only if, it is seen as the most justifiable action towards that goal state that is available within the constraints of reality.'%
\footnote{
\citet[p.\ 255]{Csibra:1998cx}; cf.\ \citet{Csibra:2003jv}.
A related but different `principle of efficiency' has been formulated by \citet[p.\ 1061]{Southgate:2008el}:
`goal attribution requires that agents expend the least possible amount of energy within their motor constraints to achieve a certain end.'
}
\end{quote}
%
Taking this as a rough starting point,
we propose that
goal ascription can involve 
identifying which potentially desirable or beneficial outcomes 
an observed action is the best available means to achieving.
More carefully, our proposal is that:
%
\begin{idescription}
\item[(E$_1$)] actions of type $a$ are normally a means of realising outcome $G$ in situations with the salient features of this situation;
\end{idescription}
%
and:
%
\begin{idescription}
\item[(E$_{2}$)] no alternative type of action, $a'$, is available to the agent and such that actions of this type would be normally be better\footnotemark \ means of realising outcome $G$ in situations which have the salient features of this situation;
\footnotetext{
An action of type $a'$ is a better means of realising outcome $G$ in a given situation than an action of type $a$ if, for instance, actions of type $a'$ normally involve less effort than actions of type $a$ 
in situations with the salient features of this situation 
and everything else is equal; 
or if, for example, actions of type $a'$ are normally more likely to realise outcome $G$ than actions of type $a$
in situations with the salient features of this situation 
and everything else is equal.
}
\end{idescription}
%
and:
%
\begin{idescription}
\item[(E$_3$)] the occurrence of outcome $G$ is potentially beneficial for the agent or typically desirable to agents of this type,
and there is no other outcome, $G'$, which would be more beneficial to the agent or typically more desirable to agents of this type
where  (E$_1$) and (E$_2$) both hold of $G'$ and $a$
\end{idescription}
%
jointly constitute defeasible evidence for the conclusion that:
%
\begin{idescription}
\item[(C)] $G$ is a goal to which action $a$ is directed.
\end{idescription}
%
We further suggest that the above inference from (E$_1$)-(E$_3$) to (C) is a route to knowledge of the goals of actions in this sense:
it would be possible to know the premises without already knowing the conclusion
and 
in some cases knowing the premises might put one in a position to know the conclusion.

This proposal might be extended in various ways.

Fact that this route to knowledge exists shows that goal ascription doesn't depend on mental state ascription.

Don't suggest this is the only route to knowledge.
Not claiming to be offering a comprehensive theory of the evidential basis of goal ascription.

@@

***Return to the leading theories?
We do not take this to show that the leading theories must be entirely discarded.
If we are right about interaction making evidence available, 
either leading theory can be extended accordingly.
***

To show that this route to knowledge is potentially valuable,
we argued that it enables interpreters to overcome two problems,
the problem of opaque means
and the problem of false belief.
These problems may rarely arise for human adults thanks to sophisticated linguistic communication and extensive knowledge of how things can be achieved.  
But interest in the evolution or development of mindreading
motivates focusing on 
interpreters with
  absent or fledgling communicative skills,
  limited insight into others' minds
  and narrow knowledge of how things work.
For such interpreters
the problems of false belief and opaque means may easily arise,
particularly in cases involving novel tools 
and 
communicative gestures.
Interpreters 
could overcome the two problems 
by exploiting the your-goal-is-my-goal route to knowledge.
It follows that 
expertise with 
  tools 
  and
  communicative gestures
does not presuppose rich insights into others' intentions and beliefs.
Such expertise,
far from  presupposing sophisticated forms of mindreading,
might instead
play a role in explaining their emergence in evolution or development (or both).

>>>>

Because goal ascription has received little attention,
in this section
we shall 
briefly review some potential benefits of being able to identify goals
and then 
consider what sort of evidence might support goal ascription.

It is a familiar idea that goal ascription enables one to learn from others' successes.
For example,
if you know or can guess that another agent's actions are directed to opening a nut,
you may then be in a position to infer that the unfamiliar pattern of actions she is performing constitute a means to open nuts.
%\citep{Horner:2005pj}
In a different case, knowing the goal of another agent's actions may enable you to discover a new use for a tool.
A slightly less familiar idea is that goal ascription enables one to learn from others' failures as well as their successes.
For example, suppose that while you are searching for some peanuts 
another agent attempts but fails to reach for a closed container.
In some circumstances,
if you know that the goal of the agent's action was to obtain the peanuts
then you now have evidence as to where they might be.\footnote{
\citet{hare_chimpanzees_2004} exploit this principle in testing chimpanzees' abilities to ascribe goals.
}
This is one illustration of how goal ascription could in principle enable us to learn from others' failures 
(\citealp{Want:2001hp} offer another).

@@

Now that we have reviewed some of the benefits goal ascription can bring,
what evidence could support ascriptions of goals to actions?
Consider the claim that 
knowledgeably identifying the goals to which actions are directed
depends on 
knowledge of the agents' intentions
which in turn depends 
on knowledge of their beliefs, desires and other mental states.
If this claim were true,
there would be no such thing as a theory of the evidential basis of goal ascription,
 only a theory of the evidential basis of mindreading more generally.

The falsity of this claim seems to be assumed in both developmental and comparative research on goal ascription.%
\footnote{ 
Compare \citet{Gergely:1995sq},
	\citet{Woodward:1998dm} and
	\citet{Penn:2007ey}
among many others.
}
We know of no compelling reason for taking this claim to be true.
After all,
it is consistent with rejecting the claim to recognize that
 ascriptions of goals to actions
 constrain, and are constrained by,
 ascriptions of intentions and other mental states.
The existence of such constraints does not show that one could 
not know something about the goals of actions 
while  knowing nothing about the agents' mental states.
Further,
there do seem to be situations where knowledge of agents' goals does not require any knowledge of their mental states.
For instance,
suppose that two people are sitting opposite each other at a low table
 which is 
sparsely populated with objects.
The objects are all out in the open; manifestly, both can clearly see them.
If one person reaches to grasp one of these objects (the duck, say), 
must the other ascribe beliefs or other mental states in order to knowledgeably identify the goal of her action as that of grasping the duck?
On the face of it, she need not.  
Even if she had no ability to ascribe mental states, it seems she might nevertheless be in a position to identify the goal of the other's action.

While not decisive,
these considerations are perhaps sufficient to motivate 
exploring what evidence might ground goal ascription.
According to 
what Csibra and Gergely call `the principle of rational action':
%
\begin{quote}
`an action can be explained by a goal state if, and only if, it is seen as the most justifiable action towards that goal state that is available within the constraints of reality.'%
\footnote{
\citet[p.\ 255]{Csibra:1998cx}; cf.\ \citet{Csibra:2003jv}.
A related but different `principle of efficiency' has been formulated by \citet[p.\ 1061]{Southgate:2008el}:
`goal attribution requires that agents expend the least possible amount of energy within their motor constraints to achieve a certain end.'
}
\end{quote}
%
Taking this idea as a rough starting point,
we propose that
%goal ascription could sometimes involve 
%identifying which potentially desirable outcomes 
%an observed action is the best available means to achieving.
%More carefully, our proposal is that 
these facts:
%
\begin{enumerate}
%
\item action $a$ is directed to some goal;
%
\item actions of $a$'s type are normally capable of being means of realising outcomes of $G$'s type in situations with the salient (to any concerned) features of this situation;
% 
\item no alternative type of action is both 
typically available to agents of this type 
and also 
such that actions of this type would be normally be significantly better\footnotemark \ means of realising outcome $G$ in situations with the salient features of this situation;
\footnotetext{
An action of type $a'$ is a better means of realising outcome $G$ in a given situation than an action of type $a$ if, for instance, actions of type $a'$ normally involve less effort than actions of type $a$ 
in situations with the salient features of this situation 
and everything else is equal; 
or if, for example, actions of type $a'$ are normally more likely to realise outcome $G$ than actions of type $a$
in situations with the salient features of this situation 
and everything else is equal.
}
%
\item the occurrence of outcome $G$ is typically desirable for agents of this type;
%
\end{enumerate}
%
and
%
\begin{enumerate}[resume]
\item there is no other outcome, $G'$, 
the occurrence of which would be at least comparably desirable for agents of this type 
and where (2) and (3) both hold of $G'$ and $a$
%
%where
%	the occurrence of $G'$ would be at least comparably desirable for agents of this type 
%%where
%%	$G$'s occurrence is not a means to $G'$'s nor conversely,
%and
%where  
%	(2) and (3) both hold of $G'$ and $a$
\end{enumerate}
%
may jointly constitute defeasible evidence for the conclusion that:
%
\begin{enumerate}[resume]
\item $G$ is a goal to which action $a$ is directed.
\end{enumerate}
%
We suggest that the above inference, from (1)-(5) to (6), is a route to knowledge of the goals of actions in this sense:
it would be possible to know the premises without already knowing the conclusion;
and, 
in some cases knowing the premises could put one in a position to know the conclusion.%
\footnote{
Knowledge of the premises may not require explicit knowledge of what others find obvious, what is salient to them or what makes actions better for them.
It is arguably sufficient that
the individual ascribing a goal 
is entitled to 
rely on being sufficiently similar to the target of her ascription 
with respect to 
	what she finds obvious, 
	what makes one action a better or worse means of realising an outcome than another action for her,  
	what is desirable to her,
	and what is salient to her.
}

Why accept this?
Suppose that 
	(1)-(5) are facts 
and that
	$G$ is not a goal to which action $a$ is directed.
Then from (1) we know that $a$ has a goal, call it $G'$.
And from (5) we can infer that either 
(i) 
the action, $a$, was not of a type that is normally capable 
of being a means to the outcome to which it was directed,
or
(ii) 
a significantly better (for instance, a more reliable or less effortful) means of achieving $G'$ is typically available to agents of this type,
or
(iii)
$G'$'s occurrence would be significantly less desirable than $G$'s occurrence for agents of this type.
So the agent's action fell short of being appropriate or optimal in some way,
or the agent has atypical preferences or capabilities,
or the situation is not normal.
While any of these is possible---%
and perhaps in some cases even to be expected---%
there are also circumstances in which it is reasonable to suppose that none obtains,
that the situation will be normal in relevant respects,
that agents will have typical preferences and capabilities 
and that they will act in ways that are appropriate and (in the limited sense in play here) optimal.
In such circumstances, the above inference can serve as a route to knowledge of the goals of actions.
The existence of this route to knowledge shows that goal ascription doesn't invariably depend on mental state ascription.

Note that we are not suggesting that the above inference captures the only route to knowledge of the goals of actions,
or that it provides anything like a comprehensive theory of the evidential basis of goal ascription.
This route to knowledge has only limited applications.
For example, it cannot be used when an action is the best available means of achieving two or more comparably desirable outcomes.
The point of this brief discussion was only to 
consider the sort of evidence that can bear on goal ascrpition
and to 
show that 
even those incapable of ascribing mental states like belief 
might nevertheless have evidence sufficient for goal ascription.

Our narrow aim in what follows is to show that some routes to knowledge of the goals of actions 
are available only to interacting mindreaders
and not to those who merely observe.
Having in this section introduced goal ascription as a topic,
the next step is to identify an obstacle to acquiring knowledge of the goals of actions,
one which we will eventually show can be avoided where mindreaders can (or appear to be able to) interact with their targets.

@@

%***Return to the leading theories?
%We do not take this to show that the leading theories must be entirely discarded.
%If we are right about interaction making evidence available, 
%either leading theory can be extended accordingly.
%***

To show that this route to knowledge is potentially valuable,
we argued that it enables interpreters to overcome two problems,
the problem of opaque means
and the problem of false belief.
These problems may rarely arise for human adults thanks to sophisticated linguistic communication and extensive knowledge of how things can be achieved.  
But interest in the evolution or development of mindreading
motivates focusing on 
interpreters with
  absent or fledgling communicative skills,
  limited insight into others' minds
  and narrow knowledge of how things work.
For such interpreters
the problems of false belief and opaque means may easily arise,
particularly in cases involving novel tools 
and 
communicative gestures.
Interpreters 
could overcome the two problems 
by exploiting the your-goal-is-my-goal route to knowledge.
It follows that 
expertise with 
  tools 
  and
  communicative gestures
does not presuppose knowledge of others' intentions and beliefs.
Such expertise,
far from  presupposing sophisticated forms of mindreading,
might instead
play a role in explaining their emergence in evolution or development (or both).

How might this work?
***sketch the story***
Clearly this is a wild conjecture,
and not one that we have given any reason for accepting.
Our point in mentioning it is merely to illustrate how
 better understanding the role of interaction in goal ascription and mindreading generally
may eventually lead to discoveries
on how
 sophisticated forms of cognition
 emerge from
 conceptually and cognitively undemanding forms of interaction.
If the culture of tool use,
the simpler forms of communicative gesture
or the sharing of knowledge and other mental states
originates in abilities to interact,
then we need to know how interaction could lead to the emergence of these things.
Our  suggestion has been that
understanding how interacting mindreaders 
could know things about their minds which mere observers might not be in a position to know
will provide some of the details.
